---
title: "Initial Exploration Notebook"
subtitle: "CS 333 Interactive Information Visualization"
author: "Aaron, Rebecca, Connor, JJ, Winnie"
pagetitle: "Initial Exploration Notebook"
date: today

format:
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    embed-resources: true
    code-fold: false
    link-external-newwindow: true

execute:
  warning: false

from: markdown+emoji
reference-location: margin
citation-location: margin  
---

## Dataset Description

For our final project, we are thinking of using the **Macro Indicators vs Asset Returns (1985–2024)** dataset. This dataset covers 40 years of macroeconomic and financial market data across 30 countries, including both developed and emerging markets. It links macro variables such as GDP growth, inflation, policy rates, unemployment, financial stress, and exchange rate changes to returns across five asset classes: equities, bonds, real estate, commodities, and cash.

The dataset is organized across several CSV files. The two main files for our analysis are `macro_indicators.csv`, which contains annual country-level macroeconomic data, and `asset_returns.csv`, which contains annual asset returns by country. These files can be joined by country, year, region, development level, and shock event. We also have supporting files such as `macro_monthly.csv`, which provides monthly G7 macro indicators, `correlations.csv`, which contains rolling macro-asset correlations, and `recession_episodes.csv`, which identifies recession periods across countries.

## Initial Data Exploration

```{r}


library(tidyverse)


# load all datasets
asset_returns <- read_csv("data/asset_returns.csv")
macro_indicators <- read_csv("data/macro_indicators.csv")
recession_episodes <- read_csv("data/recession_episodes.csv")
correlations <- read_csv("data/correlations.csv")
macro_monthly <- read_csv("data/macro_monthly.csv") 

# quick look at the structure of the two main datasets
glimpse(asset_returns)
glimpse(macro_indicators)

# what does a typical year look like for each investment type?
asset_returns |>
  summarize(
    `Avg Stock Return` = mean(equities),
    `Avg Bond Return` = mean(bonds),
    `Avg Real Estate Return` = mean(real_estate),
    `Avg Commodity Return` = mean(commodities),
    `Avg Cash Return` = mean(cash)
  )

# how many countries and years are in the dataset?
asset_returns |> summarize(
  countries = n_distinct(country),
  years = n_distinct(year),
  total_rows = n()
)

# what crises are in the data and how many country-years were affected?
asset_returns |>
  filter(!is.na(shock_event)) |>
  count(shock_event, sort = TRUE)

```
## Data Structure

The asset_returns dataset contains 1,200 rows, one per country per year. 
The five return columns (equities, bonds, real_estate, commodities, cash) 
represent annual percentage returns for each investment type. The shock_event 
column identifies years with major economic crises such as the Global Financial 
Crisis or COVID-19.

The macro_indicators dataset contains the same 1,200 rows but captures economic 
conditions: inflation (annual price growth %), policy_rate (central bank interest 
rate %), gdp_growth (economic growth %), unemployment (jobless rate %), 
fx_change (currency movement vs USD %), and fin_stress_idx (financial stress 
score from 0-3).

## Story intro

Our three communication goals form a connected narrative. We first ask whether 
crises affect all investments equally . We then ask whether economic 
signals mean the same thing across countries . Finally we ask what 
actually drives investment performance during inflation and find that government 
policy response matters more than inflation alone.

## Communication Goals

### 1. How do macro shocks reshape cross-asset performance?

Different types of macro shocks, including growth slowdowns, inflation shocks, and periods of financial stress, may alter the relative performance of asset classes rather than affecting all assets uniformly. We want to explore whether returns across equities, bonds, real estate, commodities, and cash vary meaningfully between shock and non-shock periods, potentially creating different “winners” and “losers” depending on the macro environment. The goal is to show that asset performance depends not only on whether a crisis occurs, but also on the underlying nature of the macroeconomic shock.


The table shows that during shock years, equities drop from an average of +6.3% to -0.84%, while cash remains almost unchanged. The chart shows this pattern over time for the USA, with red shading marking crisis periods.

```{r}

shock_avg <- asset_returns |>
  mutate(shock = ifelse(is.na(shock_event), "No Shock", "Shock")) |>
  group_by(shock) |>
  summarize(
    Equities = mean(equities),
    Bonds = mean(bonds),
    `Real Estate` = mean(real_estate),
    Commodities = mean(commodities),
    Cash = mean(cash)
  )

shock_avg

# get the actual shock years for USA
usa_shocks <- asset_returns |>
  filter(country == "USA", !is.na(shock_event)) |>
  select(year, shock_event)

usa_shocks
```



```{r}
asset_returns |>
  filter(country == "USA") |>
  pivot_longer(cols = c(equities, bonds, real_estate, commodities, cash),
               names_to = "asset",
               values_to = "return") |>
  mutate(asset = str_to_title(str_replace_all(asset, "_", " ")),
         shock = ifelse(is.na(shock_event), "No Shock", "Shock")) |>
  ggplot(aes(x = year, y = return)) +
  # shade every shock year
  annotate("rect", xmin = 1986.5, xmax = 1987.5, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "red") +
  annotate("rect", xmin = 1989.5, xmax = 1991.5, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "red") +
  annotate("rect", xmin = 1996.5, xmax = 1998.5, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "red") +
  annotate("rect", xmin = 2000.5, xmax = 2002.5, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "red") +
  annotate("rect", xmin = 2007.5, xmax = 2009.5, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "red") +
  annotate("rect", xmin = 2010.5, xmax = 2011.5, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "red") +
  annotate("rect", xmin = 2014.5, xmax = 2015.5, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "red") +
  annotate("rect", xmin = 2019.5, xmax = 2023.5, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "red") +
  geom_line(color = "steelblue", linewidth = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  facet_wrap(~asset, ncol = 1, scales = "free_y") +
  labs(
    title = "All 5 Assets Respond Differently During Crises (USA)",
    subtitle = "Red shading = shock/crisis years",
    x = "Year", y = "Return (%)"
  ) +
  theme_minimal(base_size = 11) +
  theme(strip.text = element_text(face = "bold"))
```

### 2. Do macro signals mean the same thing in developed and emerging markets?

Macroeconomic variables such as inflation and policy rates may reflect fundamentally different economic conditions depending on a country’s level of development. We want to explore whether emerging markets exhibit wider ranges of inflation, interest rates, and currency movement than developed markets, indicating greater levels of macroeconomic volatility and instability. The goal is to demonstrate that the same signal — such as high inflation or elevated policy rates — can represent policy tightening in developed markets but crisis risk, inflation instability, or currency pressure in emerging markets.


The chart below compares inflation rates across two developed markets (USA, Germany) and two emerging markets (Argentina, Brazil). Argentina operates at 25-70% inflation for the entire 40-year period, while the USA rarely exceeds 5%.

```{r}
macro_indicators |>
  filter(country %in% c("USA", "Argentina", "Brazil", "Germany")) |>
  ggplot(aes(x = year, y = inflation, color = country)) +
  geom_line(linewidth = 1) +
  geom_text(data = macro_indicators |> 
              filter(country %in% c("USA", "Argentina", "Brazil", "Germany"), 
                     year == 2024),
            aes(label = country), 
            hjust = -0.1, size = 3.5) +
  scale_x_continuous(limits = c(1985, 2028)) +
  scale_y_continuous(labels = scales::percent_format(scale = 1, suffix = "%")) +
  labs(
    title = "Inflation Rates Are Far Higher and More Volatile in Emerging Markets\n Than Developed Markets",
    x = "Year",
    y = "Annual Inflation Rate (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```



### 3. Does inflation alone determine asset performance, or does policy response matter more?

Asset returns during inflationary periods may depend not only on the level of inflation itself, but also on how central banks respond through monetary policy. We want to explore whether traditional hedges such as bonds continue to provide protection during periods of rising inflation and increasing policy rates. The goal is to show that macroeconomic regimes are shaped by the interaction between inflation and policy response, rather than by a single variable in isolation.



We split developed market country-years into inflation buckets and compare bond returns depending on whether policy rates were high or low. The key finding is in the 5-10% inflation bucket, bonds lose money when rates are kept low, but hold up when central banks respond aggressively.


```{r}

# join asset returns and macro indicators together
joined <- asset_returns |>
  inner_join(macro_indicators, by = c("country", "year", "dev_level", "region"))

joined |>
  mutate(
    rate_level = ifelse(policy_rate > median(policy_rate, na.rm = TRUE),
                        "High Policy Rate", "Low Policy Rate"),
    inflation_bucket = cut(inflation, 
                           breaks = c(-Inf, 2, 5, 10, Inf),
                           labels = c("Under 2%", "2-5%", "5-10%", "Over 10%"))
  ) |>
  filter(dev_level == "developed") |>
  group_by(inflation_bucket, rate_level) |>
  summarize(avg_bond = mean(bonds, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = inflation_bucket, y = avg_bond, fill = rate_level)) +
  geom_col(position = "dodge") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_fill_manual(values = c("High Policy Rate" = "#E05C5C",
                                "Low Policy Rate" = "#4A90D9")) +
  scale_y_continuous(labels = scales::percent_format(scale = 1, suffix = "%")) +
  labs(
    title = "Bond Returns Fall When Central Banks Keep Rates Low During High Inflation",
    x = "Inflation Level",
    y = "Avg Bond Return (%)",
    fill = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```